---
title: Note 24
subtitle: group operations
author: Byeong-Hak Choe
institute: SUNY Geneseo
date: 2024-05-07
format: 
  html
    # toc: true
    # toc-depth: 2
    # toc-expand: true
    # toc-title: Contents
# code-fold: true
code-tools: true

execute:
  echo: true
  eval: true
  message: false
  warning: false
  fig-width: 9
  fig-height: 6
---
```{r}
#| include: false
library(tidyverse)
library(skimr)
library(broom)
library(hrbrthemes)
library(nycflights13)
theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5) ),
                axis.title.y = element_text(size = rel(1.5) ))
          )

```




# group operations

```{r}
not_cancelled <- flights |> 
  filter(!(is.na(dep_delay) | is.na(arr_delay)))
```

Q. Describe the distribution of a plane’s average arrival delay. 

- Airplanes are identified by their tail number (tailnum)

```{r}
by_tailnum <- not_cancelled |> 
  group_by(tailnum) |> 
  summarise(mean_delay = mean(dep_delay, na.rm = T),
            n = n())
```

```{r}
ggplot(by_tailnum, aes(x = mean_delay)) +
  geom_histogram(bins = 100)
```



Q. How long is the longest average arrival delay?
```{r}
by_tailnum <- by_tailnum |> 
  arrange(-mean_delay)
```

Q. Describe the relationship between (1) the average distance from airports in New York City to a destination airport and (2) the average arrival dealy experienced on flights to that specific airport.


```{r}
by_dest <- not_cancelled |> 
  group_by(dest) |> 
  summarise(n = n(),
            dist = mean(distance, na.rm = T),
            delay = mean(arr_delay, na.rm = T))
```


```{r}
ggplot(by_dest, aes(x = dist, y = delay)) +
  geom_point(alpha = .4) +
  geom_smooth()
```



```{r}
by_dest2 <- by_dest |> 
  filter(dist <= 3000)

ggplot(by_dest2, aes(x = dist, y = delay)) +
  geom_point(alpha = .4) +
  geom_smooth()
```




```{r}
by_dest3 <- by_dest |> 
  filter(dist <= 3000, delay > 0)

ggplot(by_dest3, aes(x = dist, y = delay)) +
  geom_point(alpha = .4) +
  geom_smooth()
```








```{r}
by_dest3 <- by_dest |> 
  filter(dist <= 3000, n > 500)

ggplot(by_dest3, aes(x = dist, y = delay)) +
  geom_point(alpha = .4, aes(size = n)) +
  geom_smooth()
```

Q. Calculate the average arrival delay for each date.
Q. Given that the arrival delay is positive, calculate the average arrival delay for each date.
```{r}

by_day <- not_cancelled |> 
  group_by(year, month, day) |> 
  summarize(avg_delay = mean(arr_delay, na.rm = T),
            avg_delay_pos = mean(arr_delay[arr_delay > 0], na.rm = T),
            )

```



```{r}
by_dest_sd <- not_cancelled |> 
  group_by(dest) |> 
  summarize(dist_sd = sd(distance)) |> 
  arrange(-dist_sd)
```


Q. When do the first and last flights leave each day?

```{r}
first_last <- not_cancelled |> 
  group_by(year, month, day) |> 
  summarize(first = min(dep_time, na.rm = T),
            last = max(dep_time, na.rm = T)
            )
```



# Classwork 12

```{r}
beer_markets <- read_csv('https://bcdanl.github.io/data/beer_markets_all.csv')
```







