---
title: Lecture 9 - Note
subtitle: Data Visualization - Aesthetic Mappings and Facets
author: 
  - name: Byeong-Hak
date: 2024-02-22

execute: 
  echo: true    # true false
  eval: true    # true false
  warning: false    # true false
  message: false    # true false
  fig-width: 9
  # fig-height: 5
  
format: 
  html:
    toc: true    # true false
    number-sections: true    # true false
    code-fold: false    # true false show       # https://quarto.org/docs/output-formats/html-code.html
    code-tools: true    # true false
    highlight-style: atom-one        # atom-one tango espresso   # https://quarto.org/docs/output-formats/html-code.html#highlighting
---


```{r setup}
#| include: false
library(tidyverse)
library(hrbrthemes)
library(skimr)
library(DT)

```


# Part 3. R Basics

## Question 0.
Provide your GitHub username.

*Answer*:

bclog1


<br><br>


## Question 1.
base-R provides the R object `state.name`. Write an R code to assign `state.name` to a variable, `US_states`.

*Answer*:
```{r}
US_states <- state.name
US_states
```

<br><br>

## Question 2.
The `temp_F` vector contains the average high temperatures in January for the following cities: Seoul, Lagos, Paris, Rio de Janeiro, San Juan, and Rochester.

```{r}
temp_F <- c(35, 88, 42, 84, 81, 30)
```

Create a new vector named `temp_C` that stores the converted Celsius temperatures. Below is the conversion formula:

$$
C = \frac{5}{9}\times(F - 32)
$$

*Answer*:
```{r}
temp_C <- (5/9) * (temp_F - 32)
temp_C
```

<br><br>


## Question 3.
- Write an R code to calculate the standard deviation (SD) of the integer vector `x` below manually. That is to calculate the SD without using the `sd()` function.

```{r}
x <- 1:25
```

- Also, write an R code to test whether the standard deviation you calculate manually above is equal to `sd(x)`.



*Answer*:
```{r}

numerator <- x - mean(x)
numerator <- numerator^2
numerator <- sum(numerator)
numerator
denominator <- length(x) - 1
denominator

variance_x <- numerator/denominator
variance_x
sd_x_manually <- sqrt(variance_x)
sd_x_manually


sd_x <- sqrt( sum( (x - mean(x))^2 ) / (length(x) - 1) )
sd_x

sd(x)

sd_x == sd(x)
sd_x_manually == sd(x)
```

<br><br>


## Question 4.
- Write an R code to read the CSV file, `https://bcdanl.github.io/data/mlb_teams.csv` using the `tidyverse`'s `read_csv()` function, and assign it to `MLB_teams`.

*Answer*:
```{r}
pathname <- "https://bcdanl.github.io/data/mlb_teams.csv"
MLB_teams <- read_csv(pathname)

DT::datatable(MLB_teams)
```

<br><br>


## Question 5.
Write an R code to provide summary statistics---mean, standard deviation, minimum, first quartile, median, third quartile, and maximum---for variables in the `MLB_teams` data.frame.

*Answer*:
```{r}
skim(MLB_teams)
```

<br><br><br>




# ggplot discussion

```{r}
df_mpg <- mpg
?mpg

```

