---
title: Homework 1
subtitle: Survey, Personal Website, and R Basics
date: last-modified
from: markdown+emoji
comments: false
---

# Direction

- Please submit your **Quarto document for Part 3 in Homework 1** to the Brightspace with the name below:

  - `danl-200-hw1-LASTNAME-FIRSTNAME.qmd`\
  ( e.g., `danl-200-hw1-choe-byeonghak.qmd` )

- The due is February 20, 2024, 2:00 P.M.

- Please send Byeong-Hak an email (`bchoe@geneseo.edu`) if you have any questions.

<br><br><br>


# Part 1. Survey

Please complete the following survey:

- [Link to Google Form Survey](https://docs.google.com/forms/d/e/1FAIpQLSd9_zJ3lzcbcGpEPCWCSBL7q7ZSRX5Gxkot-fIaQQupPmCcSg/viewform?usp=sf_link)

<br><br><br>


# Part 2. Quarto Website

1. Decorate your website:
  - Replace `YOUR NAME` with your name in `_quarto.yml` and `index.qmd`.
  - Describe yourself in `index.qmd`.
  - Add the picture file (e.g., `png`) of your profile photo to `img` directory. Then correct `img/profile.png` in `index.qmd` accordingly.
  - Add the PDF file of your resumé to the website working directory in your laptop (or Posit Cloud).
  - Correct links for your resumé, LinkedIn, email, and optionally social media.
  - Make sure that you do not have any broken links in your website.

2. Add a link to the page of "R Basics" to the navigation bar using `danl-200-quarto.qmd`.

3. Use the 3-step git commands (`git add`, `git commit`, and `git push`) to update your online website.


<br><br><br>


# Part 3. R Basics

## Question 0.
Provide your GitHub username.

*Answer*:
```{r}

```

<br><br>


## Question 1.
base-R provides the R object `state.name`. Write an R code to assign `state.name` to a variable, `US_states`.

*Answer*:
```{r}

```

<br><br>

## Question 2.
The `temp_F` vector contains the average high temperatures in January for the following cities: Seoul, Lagos, Paris, Rio de Janeiro, San Juan, and Rochester.

```{r}
temp_F <- c(35, 88, 42, 84, 81, 30)
```

Create a new vector named `temp_C` that stores the converted Celsius temperatures. Below is the conversion formula:

$$
C = \frac{5}{9}\times(F - 32)
$$

*Answer*:
```{r}
temp_F <- c(35, 88, 42, 84, 81, 30)

```

<br><br>


## Question 3.
- Write an R code to calculate the standard deviation (SD) of an integer vector `x` manually. That is to calculate the SD without using the `sd()` function.

```{r}
x <- 1:25
```

- Write an R code to test whether the standard deviation you calculate manually above is equal to `sd(x)`.



*Answer*:
```{r}


```

<br><br>


## Question 4.
- Write an R code to read the CSV file, `https://bcdanl.github.io/data/mlb_teams.csv` using the `tidyverse`'s `read_csv()` function, and assign it to `MLB_teams`.

*Answer*:
```{r}

```

<br><br>


## Question 5.
Write an R code to provide summary statistics---mean, standard deviation, minimum, first quartile, median, third quartile, and maximum---for variables in the `MLB_teams` data.frame.

*Answer*:
```{r}

```

<br><br><br>


