---
title: Homework 3
subtitle: Data Visualization Basics
date: last-modified
from: markdown+emoji
comments: false
code-fold: true

execute: 
  warning: false
  message: false
  fig-width: 9
  fig-height: 7
  fig-align: center
---

```{r}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(DT)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 90,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

# Direction

- Download the Homework 3 Quarto document, `danl-200-hw-03-q.qmd`, by clicking [this link](https://drive.google.com/file/d/17LjxtbKhnjz_YsxGtcb-nLsqNTC_q_rd/view?usp=share_link).

- Provide both your answer to each question in **Homework 3** in the `danl-200-hw-03-q.qmd`.

- Please submit your **Quarto document for Homework 3** to the Brightspace with the name below:

  - `danl-200-hw3-LASTNAME-FIRSTNAME.qmd`\
  ( e.g., `danl-200-hw3-choe-byeonghak.qmd` )

- The due is March 27, 2024, 11:59 P.M.

- Please send an email to Byeong-Hak (`bchoe@geneseo.edu`) if you have any questions or clarifications.

<br><br><br>



# Customer Data with Health Insurance

Read the data file, `custdata_rev.csv`, as the data.frame object with the name, `health_cust`:

```{r}
#| echo: true
#| code-fold: false
library(tidyverse)
library(skimr)

health_cust <- read_csv(
  'https://bcdanl.github.io/data/custdata_rev.csv'
)
```


```{r}
#| echo: false
#| eval: true

rmarkdown::paged_table(health_cust)
```

<br>

## Variable Description for `health_cust`

The following describes the variables  in the `health_cust` data.frame.

- `custid`: ID of customer
- `sex`: Sex
- `is_employed`: Employment status
  - `NA`: Unknown or not applicable
  - `TRUE`: Employed
  - `FALSE`: Unemployed
- `income`: Income (in $)
- `marital_status`: Marital status
- `housing_type`: Housing type
- `recent_move`: 
  - `TRUE`: Recently moved
  - `FALSE`: Not recently moved
- `age`: Age
- `state_of_res`: State of residence
- `gas_usage`: Gas usage
  - `NA`: Unknown or not applicable
  - `001`: Included in rent or condo fee
  - `002`: Included in electricity payment
  - `003`: No charge or gas not used
  - `004`-`999`: $4 to $999 (rounded and top-coded)
- `health_ins`: Health insuarance status
  - `TRUE`: customer with health insuarance 
  - `FALSE`: customer without health insuarance 

<br><br>

## Question 1
Report summary statistics including a mean value 
for each logical variable in `health_cust`.

**Answer:**

```{r}

```


```{r}
#| echo: false
#| eval: true


```


<br>





## Question 2
Use `health_cust` data.frame to create a new data.frame, `health_cust2`, in which 
(1) `age` is greater than 0 and less than 100; and
(2) `income` is positive.

**Answer:**

```{r}

```


<br>





## Question 3 
Provide ggplot code and comments to describe 
the distribution of `income`.

**Answer:**

```{r}

```


<br>





## Question 4 
Provide ggplot code and comments to describe 
the distribution of `marital_status`.

**Answer:**

```{r}

```

- Provide some comments

<br>





## Question 5 
Suppose you randomly pick one customer's observation 
from each state in data.frame `health_cust`.

Use ggplot code to find:
(1) in which state a customer is most likely to have a health insurance; and
(2) in which state a customer is least likely to have a health insurance.

**Answer:**

```{r}
#| fig-width: 12
#| fig-height: 18

```

<br>





## Question 6 
Provide ggplot code and comments to describe 
the relationship between `age` and `income`.

**Answer:**

```{r}
```


<br>





## Question 7 
Provide ggplot code and comments to describe 
how the distribution of `age` varies by `health_ins`.

**Answer:**

```{r}


```

<br>





## Question 8 
Provide ggplot code and comments to describe 
how the distribution of `marital_status` varies by `health_ins`.

**Answer:**

```{r}


```

<br>





## Question 9 
Provide ggplot code and comments to describe how the distribution of `marital_status` varies by `housing_type`.

**Answer:**

```{r}


```

<br>



