---
title: Classwork 11
subtitle: Summarize `data.frame` with `summarize()`
date: 2024-04-23
execute: 
  echo: true
  warning: false
  message: false
  
from: markdown+emoji
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(nycflights13)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```



# Question 1. NYC dogs

- For Question 1, run the following R command to read the `nyc_dogs_cleaned.csv` file.

```{r}
#| eval: true
#| echo: true
nyc_dogs <- read_csv('https://bcdanl.github.io/data/nyc_dogs_cleaned.csv')
```


```{r}
#| echo: false
#| eval: true
#| results: asis
rmarkdown::paged_table(nyc_dogs) 
```


<br>


## Q1a

- Find the five most popular `breed`s.

*Answer*:
```{r}

```

<br><br>

## Q1b

- Provide both (1) ggplot code and (2) a simple comment to describe how the distribution of the five popular `breed`s varies by `borough`. 

*Answer*:
```{r}

```

<br><br>


## Q1c

- Find the five most popular `breed`s in each `borough`.


*Answer*:
```{r}

```

<br><br>


## Q1d

- Find the five most popular dog `name`s for each `gender`.


*Answer*:
```{r}

```

<br><br>


## Q1e

- Find the five most popular dog `name`s for each `gender` in each `borough`.


*Answer*:
```{r}

```

<br><br>


## Q1f

- Assume that all dogs in the `nyc_dogs` data frame are alive as of today.
  - Provide both (1) ggplot code and (2) a simple comment to describe the distribution of age for each `borough`.


*Answer*:
```{r}

```

<br><br>

