---
title: Classwork 12
subtitle: Data Transformation
date: 2024-05-02
execute: 
  echo: true
  warning: false
  message: false
  
from: markdown+emoji
---
```{r}
#| include: false

library(knitr)
library(rmarkdown)
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(nycflights13)

theme_set(theme_ipsum() +
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = element_text(size = rel(1.5)),
                axis.title.y = element_text(size = rel(1.5)),
                legend.title = element_text(size=rel(1.25))
                ))
```



# Part 1 - Beer Markets

```{r}
beer_markets <- read_csv('https://bcdanl.github.io/data/beer_markets_all.csv')
```


```{r}
#| echo: false
#| eval: true
#| results: asis
rmarkdown::paged_table(beer_markets) 
```

<br><br>


## Question 1
Use skimr::skim() to get summary statistics (e.g, mean, standard deviation, median, minimum, maximum) about the price for one bottle of beer (12 oz) for each `brand` when there was promo and when there was not.

*Answer*:
```{r}

```

<br><br>


## Question 2

- For each household, calculate the number of beer transactions.
- For each household, calculate the proportion of each beer brand choice.

*Answer*:
```{r}

```

<br><br>

## Question 3
- Among households that purchased BUD LIGHT at least once, what proportion only bought BUD LIGHT?

- Among households that purchased BUSCH LIGHT at least once, what proportion only bought BUSCH LIGHT?

- Among households that purchased COORS LIGHT at least once, what proportion only bought COORS LIGHT?

- Among households that purchased MILLER LITE at least once, what proportion only bought MILLER LITE?

- Among households that purchased NATURAL LIGHT at least once, what proportion only bought NATURAL LIGHT?

- Which beer brand does have the largest proportion of such loyal consumers?


*Answer*:
```{r}

```

<br><br>





# Part 2 - NYC Housing Markets

```{r}
nyc_housing_sales <- read_csv('https://bcdanl.github.io/data/nyc_housing_sales_2006-2023.csv')
```


```{r}
#| echo: false
#| eval: true
#| results: asis
rmarkdown::paged_table(nyc_housing_sales) 
```

<br><br>

## Question 4
Use skimr::skim() to get summary statistics (e.g, mean, standard deviation, median, minimum, maximum) about `sale_price` for each `borough`.


*Answer*:
```{r}

```

<br><br>


## Question 5
Add the new variable, `age`, to data.frame, nyc_housing_sales, using the `year` and `year_built` variables.

*Answer*:
```{r}

```

<br><br>


## Question 6
Provide both (1) ggplot code and (2) a simple comment to describe how the daily trend of `sale_price` varies by `borough`, on average.


*Answer*:
```{r}

```

<br><br>



## Question 7
Identify the neighborhood with the highest average housing sales price for each month of each year.

*Answer*:
```{r}

```

<br><br>
