---
title: Classwork 5
subtitle: R Basics
date: 2023-02-08
from: markdown+emoji
---


# Direction
- Open the Quarto document, `danl-200-quarto.qmd`.
- Use `danl-200-quarto.qmd` as a template to answer the following questions.

- Shortcut for creating a code chunk
  - Windows: **Alt + Ctrl + I**
  - Mac: **option + command + I**

- To run a code code chunk:
  - Click the play button in a code chunk at the top right corner.
  - Shortcut:
    - Windows: **Ctrl + Shift + Enter**
    - Mac: **command + shift + return**


- To run a current line of code in a code chunk:
  - Shortcut:
    - Windows: **Ctrl + Enter**
    - Mac: **command + return**



# Question 1
In R, what are the different data types and data structures we discuss in our course? Provide an example of each.

*Answer*
```{r}

```

<br><br>


# Question 2

```{r}
fare <- "10.00"
tip <- "2.00"
tax <- "0.80"
```

Write an R code that uses `str_c()` function to print out the following message: 

<div style="font-family:courier">The total trip cost is: $12.80</div>
  
*Answer*
```{r}

```


<br><br>


# Question 3
- Read the following CSV file as an object, `tips`.
  - `https://bcdanl.github.io/data/tips_seaborn.csv`

*Answer*
```{r}

```


<br><br>


# Question 4
- Use `DT::datatable()` for the `tips` data.frame from Question 3.

*Answer*
```{r}

```


<br><br>


# Question 5
- Use `skimr::skim()` for the `tips` data.frame from Question 3.

*Answer*
```{r}

```

<br><br>


# Discussion 


<font size = "5">Welcome to our Classwork 5 Discussion Board! 👋 </font>


This space is designed for you to engage with your classmates about the material covered in Classwork 5.

Whether you are looking to delve deeper into the slides, share insights, or have questions about the content, this is the perfect place for you.

If you have any specific questions for Byeong-Hak (**@bcdanl**) regarding the Classwork 5 materials or need clarification on any points, don't hesitate to ask here. 

Let's collaborate and learn from each other!

